---
title: "APEX Variable Documentation"
author: "Tessa LiVoti"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    number_sections: true
    highlight: kate
    cards: true
    css: apex.css
    toc_float: 
      toc_collapsed: TRUE
    toc_depth: 2
---


<style type="text/css">
.main-container {
  max-width: 1000px !important;
  margin-left: auto;
  margin-right: auto;
}
div.rect {
  border: 2px solid #035C79;
  border-radius: 15px;
  padding: 10px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(RPostgreSQL)
library(stringr)
library(dplyr)
library(readxl)
library(knitr)
library(readxl)
library(DT)

# database login info taken out for security - contact me directly if you have questions 

dartnet_snomed <- read_excel("code_lists/snomed_apex_valuesets.xlsx")
repo <- dbGetQuery(db, 'select * from repository.repository')
value_set <- dbGetQuery(db, 'select * from repository.value_set')

drug_cat <- read_excel('code_lists/all_drug_cat.xlsx')


```


# COPD Population Selection

## Overall {.tabset .tabset-fade .tabset-pills}

### Explanation {-}

The overall COPD population is selected based on meeting the following criteria:

* Patients who have had an appointment in the last 2 years AND either:
    + Have been diagnosed/re-coded with COPD in the last 2 years OR 
    + Have had a prescription for a COPD inhaler in the last 2 years AND 
    + have a COPD diagnosis ever


## Asthma Population {#asthma .tabset .tabset-fade .tabset-pills}

### Explanation {-}

The final asthma  population is selected based on meeting <b>all</b> of the following criteria:

* Asthma diagnosis code (either ICD or SNOMED)



### Asthma Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3> Asthma Diagnosis Codes</h3></center>
```{r qofAsthma_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}


Asthma_codes <- filter(repo, value_set_id == 305) %>%
  select(c(code, code_system, description)) 

colnames(Asthma_codes) <- c('Code', 'Code System', 'Description')
#qofAsthma_codes$Term <- gsub('[X]', '', qofAsthma_codes$Term, fixed = T)

datatable(Asthma_codes, escape = F, filter = 'top', rownames = F, extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## COPD Diagnosis Codes {#copd .tabset .tabset-fade .tabset-pills}

### Explanation {-}

### COPD Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>COPD Diagnosis Codes</h3></center>
```{r qofcopd_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

COPD_codes <- filter(repo, value_set_id == 127) %>%
  select(c(code, code_system, description))


colnames(COPD_codes) <- c('Code', 'Code System', 'Description')

datatable(COPD_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>



# Demographics 

## Age {#age}

Age is calculated by subtracting patient's birth year from the year of the audit.

## Gender {#gender}

Gender is pulled from the patient's record.

## Weight {#weight}

Weight is stored as pounds or ounces in the patient's EHR and converted to kg using either: 

* weight_kg = pounds/2.205
* weight_kg = oz/35.274

## Height {#height}

Height is stored in inches in the patient's EHR and converted to meters using: 

* height_m = inches/39.37

## Body Mass Index (BMI) {#bmi}

BMI is calculated using height (in m) and weight (in kg) using: 

* bmi = weight_kg/(height_m^2)

It is then categorized into one of the following: 

* under weight: BMI < 18.5
* normal weight: 18.5 =< BMI < 25
* over weight: 25 =< BMI < 30
* obese: 30 =< BMI 

## Ethnicity {#ethnicity}

Ethnicity is pulled from the patient's record.

## Race {#race}

Race is pulled from the patient's record.


# Comorbidities

## AATD {#aatd .tabset .tabset-fade .tabset-pills}

### Explanation {-}

AATD diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Bronchiectasis Diagnosis Codes</h3></center>
```{r aatd_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

aatd_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "AATD") 

colnames(aatd_codes) <- c('Code', 'Code System', 'Description')

datatable(aatd_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Acute sinusitis {#acutesin .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Acute sinusitis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Cough Diagnosis Codes</h3></center>
```{r acute_sinusitis_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

acute_sinusitis_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "acute sinusitis")

colnames(acute_sinusitis_codes) <- c('Code', 'Code System', 'Term')

datatable(acute_sinusitis_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Anemia {#anemia .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Anemia diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Cystic Fibrosis Diagnosis Codes</h3></center>
```{r anemia_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

anemia_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Anemia")


colnames(anemia_codes) <- c('Code', 'Code System', 'Term')

datatable(anemia_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Ankle Edema  {#ankle .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Ankle Edema  diagnosis code ever.


### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Anxiety Diagnosis Codes</h3></center>
```{r ankle_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

ankle_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Ankle Edema")


colnames(ankle_codes) <- c('Code', 'Code System', 'Term')

datatable(ankle_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Anxiety  {#anx .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Anxiety diagnosis code ever.


### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Anxiety Diagnosis Codes</h3></center>
```{r anx_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

anx_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Anxiety")


colnames(anx_codes) <- c('Code', 'Code System', 'Term')

datatable(anx_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Chronic Sinusitis  {#chronicsin .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Chronic Sinusitis diagnosis code ever.


### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Anxiety Diagnosis Codes</h3></center>
```{r chronicsin_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

chronicsin_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "chronic sinusitis")

colnames(chronicsin_codes) <- c('Code', 'Code System', 'Term')

datatable(chronicsin_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Diabetes  {#diabetes .tabset .tabset-fade .tabset-pills}

### Explanation {-}

To be categorized with diabetes, patients must meet <b>all</b> of the following criteria:

* Diabetes diagnosis code ever, regardless of type I or II
* No diabetes resolved codes since their last diabetes diagnosis

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Diabetes Diagnosis Codes</h3></center>
```{r diabetes_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

diabetes_codes <- filter(repo, value_set_id == 289) %>%
  select(c(code, code_system, description))


colnames(diabetes_codes) <-  c('Code', 'Code System', 'Term')

datatable(diabetes_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Eczema {#eczema .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Eczema diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Eczema Diagnosis Codes</h3></center>
```{r eczema_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

eczema_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Eczema")


colnames(eczema_codes) <- c('Code', 'Code System', 'Term')

datatable(eczema_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## GERD {#gerd .tabset .tabset-fade .tabset-pills}

### Explanation {-}

GERD diagnosis code in the audit period.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>GERD Diagnosis Codes</h3></center>
```{r GERD_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

GERD_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "GERD") 


colnames(GERD_codes) <-  c('Code', 'Code System', 'Term')

datatable(GERD_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Heart Failure {#heartfail .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Heart failure code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Heart Failure Diagnosis Codes</h3></center>
```{r hf_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

hf_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Heart failure") 


colnames(hf_codes) <-  c('Code', 'Code System', 'Term')

datatable(hf_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Hypertension {#hyperten .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Hypertension diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Hypertension Diagnosis Codes</h3></center>
```{r hyperten_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

hyperten_codes <- filter(repo, value_set_id == 306) %>%
  select(c(code, code_system, description))

colnames(hyperten_codes) <- c('Code', 'Code System', 'Term')

datatable(hyperten_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>



## Hypoxemia {#hypoxemia .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Hypoxemia diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Hypoxemia Diagnosis Codes</h3></center>
```{r hypoxemia_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

hypoxemia_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Hypoxemia")

colnames(hypoxemia_codes) <- c('Code', 'Code System', 'Term')

datatable(hypoxemia_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Lower Respiratory Tract Infection (LRTI) {#lrti .tabset .tabset-fade .tabset-pills}

### Explanation {-}

LRTI diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>LRTI Diagnosis Codes</h3></center>
```{r lrti_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

lrti_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "LRTI") 

colnames(lrti_codes) <- c('Code', 'Code System', 'Term')

datatable(lrti_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Lung Cancer  {#lungcancer .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Lung cancer code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Lung Cancer Diagnosis Codes</h3></center>
```{r lungcancer_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

lungcancer_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Lung cancer")

colnames(lungcancer_codes) <- c('Code', 'Code System', 'Term')

datatable(lungcancer_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'), 
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Metabolic Syndrome {#metasynd .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Metabolic Syndrome diagnosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Metabolic Syndrome Diagnosis Codes</h3></center>
```{r ms_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

ms_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Metabolic syndrome")


colnames(ms_codes) <- c('Code', 'Code System', 'Term')

datatable(ms_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Nasal Polyps {#nasalpolyp .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Nasal Polyp code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Nasal Polyp Diagnosis Codes</h3></center>
```{r np_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

np_codes <-  read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Nasal Polyp")

colnames(np_codes) <- c('Code', 'Code System', 'Term')

datatable(np_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Obstructive Sleep Apnea (OSA) {#osa .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Obstructive Sleep Apnea code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>OSA Diagnosis Codes</h3></center>
```{r osa_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

osa_codes <-  read_excel("code_lists/APEX_code_lists.xlsx", sheet = "osa")

colnames(osa_codes) <- c('Code', 'Code System', 'Term')

datatable(osa_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Osteoarthritis {#osteoarth .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Osteoarthritis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Osteoarthritis Diagnosis Codes</h3></center>
```{r osteoarth_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

osteoarth_codes <-  filter(repo, value_set_id == 605) %>%
  select(c(code, code_system, description))

colnames(osteoarth_codes) <- c('Code', 'Code System', 'Term')

datatable(osteoarth_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Osteoporosis {#osteopor .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Osteoporosis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Osteoporosis Diagnosis Codes</h3></center>
```{r osteopor_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

osteopor_codes <-  read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Osteoporosis")

colnames(osteopor_codes) <- c('Code', 'Code System', 'Term')

datatable(osteopor_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Pneumonia {#pneumonia .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Pneumonia code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Pneumonia Diagnosis Codes</h3></center>
```{r pneumonia_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

pneumonia_codes <-  filter(repo, value_set_id == 1040) %>%
  select(c(code, code_system, description))

colnames(pneumonia_codes) <- c('Code', 'Code System', 'Term')

datatable(pneumonia_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Rhinitis {#rhinitis .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Rhinitis code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Rhinitis Diagnosis Codes</h3></center>
```{r rhinitis_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

rhinitis_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Rhinits") 

colnames(rhinitis_codes) <- c('Code', 'Code System', 'Term')

datatable(rhinitis_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## Weight loss {#weightloss .tabset .tabset-fade .tabset-pills}

### Explanation {-}

Weight loss code ever.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Weight loss Diagnosis Codes</h3></center>
```{r weightloss_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

weightloss_codes <- read_excel("code_lists/APEX_code_lists.xlsx", sheet = "Weight loss")

colnames(weightloss_codes) <- c('Code', 'Code System', 'Term')

datatable(weightloss_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

# Pharmacotherapy

## ICS {#ics .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for ICS.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>ICS medication Codes</h3></center>
```{r ics_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

ics_codes <- filter(drug_cat, drug_type == "ics") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(ics_codes) <- c('Code', 'Term', 'Type')

datatable(ics_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## ICS/LABA {#icslaba .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for ICS/LABA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>ICS/LABA medication Codes</h3></center>
```{r ics_laba_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

ics_laba_codes <- filter(drug_cat, drug_type == "ics_laba") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(ics_laba_codes) <- c('Code', 'Term', 'Type')

datatable(ics_laba_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## ICS/LABA/LAMA {#icslabalama .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for ICS/LABA/LAMA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>ICS/LABA/LAMA medication Codes</h3></center>
```{r ics_laba_lama_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

ics_laba_lama_codes <- filter(drug_cat, drug_type == "ics_laba_lama") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(ics_laba_lama_codes) <- c('Code', 'Term', 'Type')

datatable(ics_laba_lama_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## LABA {#laba .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for LABA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>LABA medication Codes</h3></center>
```{r laba_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

laba_codes <- filter(drug_cat, drug_type == "laba") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(laba_codes) <- c('Code', 'Term', 'Type')

datatable(laba_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## LAMA {#lama .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for LAMA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>LAMA medication Codes</h3></center>
```{r lama_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

lama_codes <- filter(drug_cat, drug_type == "lama") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(lama_codes) <- c('Code', 'Term', 'Type')

datatable(lama_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## LABA/LAMA {#labalama .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for LABA/LAMA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>LABA/LAMA medication Codes</h3></center>
```{r laba_lama_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

laba_lama_codes <- filter(drug_cat, drug_type == "laba_lama") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(laba_lama_codes) <- c('Code', 'Term', 'Type')

datatable(laba_lama_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## SABA {#saba .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for SABA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>SABA medication Codes</h3></center>
```{r saba_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

saba_codes <- filter(drug_cat, drug_type == "saba") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(saba_codes) <- c('Code', 'Term', 'Type')

datatable(saba_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## SAMA {#sama .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for SAMA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>SAMA medication Codes</h3></center>
```{r sama_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

sama_codes <- filter(drug_cat, drug_type == "sama") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(sama_codes) <- c('Code', 'Term', 'Type')

datatable(sama_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


## SABA/SAMA {#sabasama .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for SABA/SAMA.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>SABA/SAMA medication Codes</h3></center>
```{r saba_sama_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

saba_sama_codes <- filter(drug_cat, drug_type == "saba_sama") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(saba_sama_codes) <- c('Code', 'Term', 'Type')

datatable(saba_sama_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## OCS {#ocs .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for OCS.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>OCS medication Codes</h3></center>
```{r ocs_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

ocs_codes <- filter(drug_cat, drug_type == "ocs") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(ocs_codes) <- c('Code', 'Term', 'Type')

datatable(ocs_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Macrolide {#macro .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for Macrolides.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Macrolide medication Codes</h3></center>
```{r macro_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

macro_codes <- filter(drug_cat, drug_type == "macrolide") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(macro_codes) <- c('Code', 'Term', 'Type')

datatable(macro_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

### Detials {-}

Only long term courses of macrolides are being included. Long term term macrolide treatment is defined as the use of macrolides that have been administered orally at least one time a week for a period of at least 3 months as referred to in [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6254503/). Short term courses for macrolides are excluded.

## PDE4 {#pde4 .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for PDE4.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>PDE4 medication Codes</h3></center>
```{r pde4_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

pde4_codes <- filter(drug_cat, drug_type == "pde4") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(pde4_codes) <- c('Code', 'Term', 'Type')

datatable(pde4_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Antibiotics {#antib .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for antibiotics.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Antibiotic medication Codes</h3></center>
```{r antib_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

antib_codes <- filter(drug_cat, drug_type == "antibiotic") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(antib_codes) <- c('Code', 'Term', 'Type')

datatable(antib_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

## Diuretics {#diuretics .tabset .tabset-fade .tabset-pills}

### Explanation {-}

RxNorm codes for diuretics.

### Diagnosis Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Diuretic medication Codes</h3></center>
```{r diur_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

diur_codes <- filter(drug_cat, drug_type == "diuretics") %>% 
  select(c(rxnorm, Generic_Name, drug_type))

colnames(diur_codes) <- c('Code', 'Term', 'Type')

datatable(diur_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>

# Risk Indicators

## Exacerbations {#exacerbations .tabset .tabset-fade .tabset-pills}

### Explanation {-}

An exacerbation is defined as any of the folloing events: 

* A code for an exacerbation found in the patient's EHR. 
* Antibiotic prescribed with evidence of a lower respiratory consultation 
* A course of oral corticosteroids with evidence of an acute respiratory infection 
* An acute course of antibiotic and oral corticosteroid at the same time

### Exacberation Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Exacerbaion Codes</h3></center>
```{r exac_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

exac_codes <- read_excel("code_lists/exacerbation_algorithm_final.xlsx", sheet = "exacerbations_codes")

colnames(exac_codes) <- c('Code', 'Term', 'Type')

datatable(exac_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


### Antibiotic Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Exacerbaion Codes</h3></center>
```{r anti_exac_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

antibiotic_codes <- read_excel("code_lists/exacerbation_algorithm_final.xlsx", sheet = "antibiotic")

colnames(exac_codes) <- c('Code', 'Term', 'Type')

datatable(exac_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


### Oral Corticosteroid Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Exacerbaion Codes</h3></center>
```{r ocs_exac_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

exac_codes <- read_excel("code_lists/exacerbation_algorithm_final.xlsx", sheet = "ocs_codes")

colnames(exac_codes) <- c('Code', 'Term', 'Type')

datatable(exac_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>


### LRTI Codes {-}

<div class = "row">
<div class = "rect", style="float:right; margin-left:5%; margin-right: 5%; width:90%;">
<center><h3>Exacerbaion Codes</h3></center>
```{r lrti_exac_codes, fig.height=4.5, dev.args = list(bg = 'transparent')}

exac_codes <- read_excel("code_lists/exacerbation_algorithm_final.xlsx", sheet = "LRTI codes")

colnames(exac_codes) <- c('Code', 'Term', 'Type')

datatable(exac_codes, escape = F, filter = 'top', rownames = F, 
          extensions = c('Buttons', 'FixedColumns'),
          options = list(dom = 'lfrtBip', buttons = c('copy', 'csv','pdf', 'print'), 
                         scrollX = T, fixedColumns = list(leftColumns = 1))) 

```
</div></div>
# References

Cui Y, Luo L, Li C, Chen P, Chen Y. Long-term macrolide treatment for the prevention of acute exacerbations in COPD: a systematic review and meta-analysis; 2018 Nov; 
Available from: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6254503/

